---
title: 'I am a MONSTR: Debuting a Wiki-based Dataset on Military Operations with Novel Strategic Technologies'
author:
  - name: J Andrés Gannon
    email: "andres.gannon@gmail.com"
    url: jandresgannon.com
    affiliation: Council on Foreign Relations
  - name: Kerry Chavéz
    email: "kerry.chavez@ttu.edu"
    url: kerrychavez.us
    affiliation: Texas Tech University
format:
  pdf:
    fig-format: png
    thanks: test text here
editor: visual
execute:
  echo: false
  warning: false
  cache: true
date: "`r format(Sys.Date(), '%B %d, %Y')`"
abstract: \singlespacing \noindent Research on strategies and force structures in modern warfare is prolific, but siloed. While some examine boots on the ground, others focus on aerial bombing or unmanned platforms. Consequently, most studies focus on the effects of one approach, seldom considering it in lieu of or conjunction with others. Furthermore, there is less knowledge on the origins and implementations of these strategic choices analyzed in isolation. The primary reason for these gaps lies with data limitations. In this paper, we introduce the first comprehensive dataset on the universe of US military interventions from 1989-2019. Scraped from the understructure of Wikipedia using SPARQL, a database query language, we offer unprecedented coverage and granularity that enables analysis of myriad determinants—domestic, international, geographic, temporal—of how states fight. The data are also nested, providing opportunities to model interventions in the strategic campaigns and operations in which they were planned. We describe and demonstrate the data’s contents and utility, then suggest how the novel, fine-grained information can contribute to theories about the effects of technological innovation on international conflict.
header-includes:
- \usepackage{tikz}
- \usepackage{pgfplots}
- \pgfplotsset{compat=newest}
- \usetikzlibrary{plotmarks}
- \usetikzlibrary{arrows.meta}
- \usepgfplotslibrary{patchplots}
- \usepackage{grffile}
- \usepackage{caption}
- \usepackage[utf8]{inputenc}
- \usepackage[doublespacing]{setspace}
- \AtBeginEnvironment{tabular}{\singlespacing}
- \usepackage{float}
- \usepackage{multirow}
- \usepackage{tablefootnote}
- \usepackage{pifont}
- \usepackage{newunicodechar}
- \usepackage{booktabs}
- \usepackage{tabularx}
- \newunicodechar{✓}{\ding{51}}
bibliography: cruisemissiles_bib.bib
---

```{r knitr_options, echo = FALSE, warning = FALSE, cache = FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
```

```{r load data}
# New data
df_interv <- readRDS(paste0(here::here(), "/data/","06a_model.rds"))

# Prior data
icb <- read.csv(paste0(here::here(),"/inst/extdata/icb2v12.csv")) %>%
  dplyr::filter(yrtrig >= 1989 &
                  actor == "USA")

imi <- read.csv(paste0(here::here(),"/inst/extdata/IMI/IMI_dataset.csv")) %>%
  dplyr::filter(start >= 19890101 &
                  intervener == 2)
imi$start <- gsub('.{4}$', '', imi$start)

mids <- read.csv(paste0(here::here(),"/inst/extdata/MIDs/MIDs_dispute-participants.csv")) %>%
  dplyr::filter(styear >= 1989 &
                  ccode == 2)

mips <- read.csv(paste0(here::here(),"/inst/extdata/MIPS/MIPS_dataset.csv")) %>%
  dplyr::filter(styr >= 1989 &
                  MP == 2)

mip <- readstata13::read.dta13(paste0(here::here(),"/inst/extdata/MIP/MIPDataKushiToft.dta")) %>%
  dplyr::filter(styear >= 1989 &
                  RemoveCase == 0)
  
prio <- read.csv(paste0(here::here(),"/inst/extdata/PRIO/UCDP-PRIO_armed_conflict_dataset.csv")) %>%
  dplyr::filter(year >= 1989 &
                  side_a_id == 3) # PRIO does not use the GW identifier. It is confirmed that this refers to the USA

rand <- read.csv(paste0(here::here(),"/inst/extdata/rand_interventionlist.csv")) %>%
  dplyr::filter(incident_start_year >= 1989)

allen_17 <- readstata13::read.dta13(paste0(here::here(), "/inst/extdata/AllenMtzMachainAerialBombingData2016CMPS.dta")) %>%
  dplyr::filter(year >= 1989,
                ccode1 == 2)

horowitz_01 <- foreign::read.dta(paste0(here::here(), "/inst/extdata/jcr01.dta")) %>%
  dplyr::filter(attacker %in% c('United States', 'Allies', 'U.S./Britain', 'NATO', 'US')) %>%
  dplyr::mutate(date = dplyr::recode(date,
                                     '1944-1945' = "1944",
                                     '1942-1944' = "1942",
                                     'Aug. 1945' = "1945",
                                     '1965-1968' = "1965",
                                     '1950-1951' = "1950")) %>%
  dplyr::mutate(date = as.numeric(date)) %>%
  dplyr::filter(date >= 1989)
```

\newpage

# Introduction

The recent interest in the means by which states fight has primarily focused on traditional forms of interstate conflict. Yet, many of the military involvements by states in recent decades, particularly great powers, has occurred not in the context of interstate wars, but military interventions that are more appropriately conceived of as "wars of choice". The actions a state takes during wars of choice are not pre-determined at the start of the intervention, rather they are revisited and consequently evolve over the course of a conflict.

Current research on how states undertake military interventions either focuses on the most aggregate level of the war or explains a particular tool of intervention in isolation. In doing so, variation in how states behave over the course of a multi-year intervention remains unexplained, as does the relationship between different campaigns and operations that constitute military interventions.

This paper remedies this problem by providing a new dataset on US military interventions at the campaign, operation, and battle level using a familiar, yet under-utilized data source: wikipedia. By simply scraping wikipedia for pages on US military interventions since 1989, our dataset identifies almost 10 times as many events as all existing datasets combined. Using wikidata and DBpedia, the two structured backends of wikipedia, we can also extract a host of politically salient covariates concerning each military operation's dates, belligerents, precise location, and military means used.

This paper proceeds as follows. Section 2 identifies our motivation, detailing the existing state of the art in understand how states intervention and the value of explaing various intervention choices and contexts. Section 3 outlines our approach in collecting new data on US military interventions from wikipedia and how the data was cleaned, organized, and validated. Section 4 produces descriptive statistics of the data to demonstrate the value-added and how scholars can use these data to augment existing studies on the actors, strategies, and locations of US military interventions. Section 5 constitutes an inductive data-driven model that predicts what military tools the US uses in an intervention given factors like the timing, location, target, incumbent president, and coalition partners. Section 6 concludes.

# Motivation

This all began with a formal model. Paper is about military strategy, which refers to "the general principles guiding the use of conventional military forces for the overall planning and execution of armed conflict" [@wallace_alliancesinstitutionaldesign_2008, 225]. We wanted to know how strong states fight, and why. Sometimes they choose losing strategies against weak actors [@mack_whybignations_1975; @arreguin-toft_howweakwin_2001]. Military myopia arguments [@ohanlon_canhightechnology_1998; @gentry_doomedfailamerica_2002; @lyall_ragemachinesexplaining_2009; @bolton_minefieldsminespacearcheology_2015], capital-rich force structure narratives [@gartzke_democracypreparationwar_2001; @caverley_mythmilitarymyopia_2009], and uncertainty in pre-war estimates of cost and success [@sullivan_waraimswar_2007] explain many of these cases. Yet sometimes, despite bureaucratic inertia, economic factor abundance, and the fog of war, advances militaries deploy ideal force structures against innovative weak enemies. We built a formal model to examine the conditions capable of explaining this variation. There it was---intuitive (in a multipage math proof kind of way), one-sided imperfect information (so perfect) carving two beautiful paths for our strong state player (insofar as Bayesian Nash equilibria are beautiful). Realizing we needed to continue in the spirit of empirical implications of theoretical models (okay, the formal model wasn't *that* good), we also realized that we had six reasonable comparative statics and nowhere to go.

We identified three barriers to testing our theory and model with existing data. First, we are interested in the planning and execution of military strategy *during* armed conflict [@wallace_alliancesinstitutionaldesign_2008, 225]. Most of the research on international conflict has focused on its beginning -- bargaining, causes, onsets \-- and end \-- duration, outcomes, aftermaths. Notwithstanding the profound knowledge they have yielded, these approaches black-box the prosecution of war. This is surprising since we think its conduct has much to tell us about its causes and consequences. Some studies breach the black box of military force to analyze approaches within, but typically focus on the possession of these platforms rather than their use, offering insight only on the latent potential for the use of force.[^1] Furthermore, the majority seem to bore a peephole into a single platform or force structure rather than pop open the box. Some look at nuclear platforms [@gartzke_determinantsnuclearforce_2014], naval power [@crisher_powerseanaval_2014], aerial bombing [@pape_bombingwinair_1996], air power more broadly [@horowitz_whendoesaerial_2001; @allen_timebombsestimating_2007; @martinezmachain_aircampaignduration_2015; @allen_understandingimpactair_2017], ballistic missiles [@reiter_ballisticmissilesinternational_2013], cruise missiles [@early_climbingladderexplaining_2022], or drones [@horowitz_whopronedrone_2022].

[^1]: Recent exceptions that focus on use include @martinezmachain_aircampaignduration_2015, @post_flyingfailcostly_2019, and @gannon_oneifland_2022.

This leads to the second barrier -- siloed scholarship that inhibits the idea of substitutability [@morgan_modelforeignpolicy_2000]. At a high level, the substantial foreign policy literature reflects this standpoint by depicting a state's toolkit in broad categories \-- military, diplomatic, economic, cultural, etc. [@brecher_frameworkresearchforeign_1969; @greenhill_coercionpowerhurt_2018]. We contend that the same logic applies within each category. Thus, leaders craft the use of military force not as isolated dichotomous decisions, but from a menu of substitutable (or combinable) options. Being domain- or platform-specific, existing datasets on the means of force are difficult to compare and harmonize. Rather than peepholes into the blackbox of military force, our research questions on the construction of strategy and force structure require it to be open and well-lit. Once again, our inquiries on how and why states fight fall between existing theory and data, highly generalized on one side, highly specific but partitioned on the other.

Finally, holding the universe of military interventions and the idea of substitutability in mind, a significant concern about assumptions of independence across observations cropped up. Hypothetically, if we seek to explain why an intervention featured solely unmanned systems, perhaps it is relevant if it is part of a broader decapitation campaign in enemy territory. Or if a sizable contingent of troops is deployed for a manpower surge in, say, a place like Iraq (just riffing here), perhaps one might expect interventions nested within such a surge campaign to feature a higher frequency of group troops. Attempting to explain force structure choices outside the strategic contexts in which they occur might lead to bias in estimating independent causal impacts. Yet existing datasets largely include interventions within the same broader operation or campaign as distinct observations. Ultimately, we desired a dataset that featured 1) measures of the means of military force across 2) a comprehensive list of open-source US military interventions that 3) captures dependence between observations using a consistent and appropriate unit of analysis. To our knowledge, it did not exist. And little wonder given how labor-intensive and dimensional the creature would be. As we weighed the task, one final barrier arose: how would we collect data of this scope and scale without using Wikipedia to *some* degree?

# Our approach

We agonized over it for some time. In the end, we realized it was right under our noses all the while. Only, it reeked of research disrepute. *Wikipedia*. We shuddered. After deliberating the pros, cons, alternatives, and tradeoffs, however, we determined that it would help us explore our research questions and make a contribution to existing scientific knowledge on military intervention. Consequently, we created a new dataset on the use of force employed in US military interventions from 1989 to 2020.

We disciplined our collection efforts definitionally, in space, and in time to curate a first cut dataset of greatest purchase.[^2] First, insofar as our case selection is most consistent with the exemplary Military Interventions by Major Powers (MIPS) dataset's definition, it merits quoting: an intervention is "a use of armed force that involves the official deployment of at least 500 regular military personnel (ground, air, or naval) to attain immediate term political objectives through action against a foreign adversary. Routine military movements and operations without a defined target like military training exercises, the routine forward deployment of military troops, non-combatant evacuation operations, and disaster relief should be excluded." [@pearson_internationalmilitaryintervention_1993, 1]. Like MIPS, we exclude threats, training exercises or mobilizations that deterrently demonstrate rather than apply force, ebbs and flows of latent manpower and materiel at permanent military bases, force engaged in diplomatic cover and evacuation, and humanitarian deployments of military units. We deviate from MIPS in that we do not set a size threshold for inclusion.

[^2]: Spoiler alert: the rest of this subsection is boring.

We also lean on the International Military Intervention (IMI) dataset's strong definition, which clarifies that events must be purposeful [@kisangani_internationalmilitaryintervention_2008]. Thus, we omit accidental and unintentionally contingent events arising in context of a planned intervention. For instance, if troops deployed along the Iraq border have an unexpected encounter with Syrians, we view it as simply that---a military intervention in Iraq that entailed a contingent clash with Syrians who meddled at the border. Being outside the politically determined mission scope, such a clash (or airspace violation) does not constitute a separate intervention. We deviate from IMI in that we do not include diplomatic cover / evacuation activities or humanitarian force.

There are other definitions and identification strategies across datasets on military force. While several cast a wider net, we are firm in our focus since we seek to understand how states fight and non-fighting activities have different antecedents and effects. We also insist that political and military objectives cannot be delinked, disqualifying nonpolitical uses of force. Table A1 in the Appendix summarizes our nearness to MIPS and IMI across definitional fine points, and our departures from RAND [@kavanagh_characteristicssuccessfulmilitary_2019] and Military Intervention Project (MIP) [@kushi_introducingmilitaryintervention_2022] which take a broader approach. Focusing on definitions of intervention, we omit other important datasets in this particular comparison that we later consider.[^3]

[^3]: Namely, these include International Crisis Behavior (ICB) which emphasizes foreign policy crises [@brecher_studycrisis_1997], Uppsala Conflict Data Program (UCDP) / Peace Research Institute Oslo (PRIO) which takes armed conflict as its unit of analysis [@gleditsch_armedconflict19462001_2002], and Militarized Interstate Disputes (MIDs) [@palmer_updatingmilitarizedinterstate_2020].

In the spirit of intellectual transparency, our dataset includes some paramilitary activities. Most notably, we capture some Central Intelligence Agency (CIA) drone strikes but not all. We defend this on two fronts. First, drones are distinct as a means of force since some are operated by the military (in combat zones) and some by the CIA (in noncombat zones). In fact, extensive criticism has been levied at the CIA drone program for its lack of military control, checks and balances, and transparency. Second, every dataset on military interventions exhibits nonrandom missingness on some drone cases due to their covert nature (even those devoted solely to drone strikes). MONSTR's coverage is appealing in that the drone strikes it does include---military and CIA---have met some threshold of public awareness to make it onto Wikipedia. Since we expect that public constraints amid the domestic politics of war are meaningful factors in how states fight, we accept this bias. In fact, given this variation we wonder what criteria determines publication on Wikipedia in the first place, a question our dataset allows us to pose and test.

Similar in logic, by nature of being collectible from the internet our cases are all open source. There are certainly interesting questions about the deployment of covert force, but we appreciate former Department of Intelligence Director Wilson's comment that "90% of intelligence comes from open sources. The other 10%, the clandestine work, is just the more dramatic. The real intelligence hero is Sherlock Holmes, not James Bond" (FPB 2010, 23). It's elementary -- we leverage the data that are available and focus on research puzzles that interface with public attention and scrutiny.

Next, we spatially confined our sample to cases featuring the United States for several reasons. Insofar as our impetus to collect these data is to examine how strong states fight, the US is a paragon. As bipolar contender since World War II and unipolar hegemon post-Cold War, its prolific intervention record also allows us to maximize observations in a pilot probe with one case. The US case is also deeply studied and data rich, allowing us to test our theories upon a bedrock and more easily compare findings. Finally, we temporally confined our scope to start in 1989, when the US first used cruise missiles in conflict. With a special interest in the politics of emerging technologies, we aimed to collect a sample with temporal consistency in the military options, from ground troops to unmanned platforms, available to conflict decision-makers. Next, we describe the data collection process, with an emphasis on justifying our source, before demonstrating the data's basic yields and potential.

## Collection

For most readers, the Wikipedia interface is rather familiar. An executive summary sits under a prominent heading, often followed by a contents box and its subsequent sections and fringed by an infobox containing common attributes and connections. For example, the "cosmic latte" page informs that this is the average hue of the universe based on spectral analysis of over 200,000 galaxies. The sections describe the discovery and naming of the color.[^4] The infobox records the color coordinates across a range of designation systems. These data, like most other sources available to scholars, are unstructured and require manual review and entry.

[^4]: Thank goodness "primordial clam chowder" didn't win.

Unlike most other sources, Wikipedia rests upon two Resource Description Format (RDF) databases: Wikidata and DBpedia [@malyshev_gettingmostout_2018]. Wikidata contains ontology of over 99 million items[^5] and their properties as "triples." The wikidata page for Germany, for example, lists a subject (Germany), a predicate (has capital), and object (Berlin). In most cases, there are dozens to hundreds of triples for each page that can be automatically extracted using SPARQL queries [@malyshev_gettingmostout_2018]. This can automate the process of identifying the universe of cases in ways that save significant time and effort compared to manual data entry. DBpedia maps items' covariates, also as triples. Since they exist in a uniform, machine-readable format, we were able to scrape a comprehensive list of public US military interventions during our time scope of analysis and their covariates using SPARQL, an RDF query language [@morsey_dbpediasparqlbenchmark_2011]. We supplemented this with DBpedia, which is a standardized and machine-readable backend to the wikipedia infoboxes. Infoboxes are defined by wikipedia as "contain important facts and statistics of a type which are common to related articles. For instance, all animals have a scientific classification (species, family and so on), as well as a conservation status. Adding an \[infobox\] to articles on animals therefore makes it easier to quickly find such information and to compare it with that of other articles." This resulted in a fairly replete dataset. For all missing data points, research assistants were assigned to complete entries from alternative sources, where possible. After reasonable efforts to locate missing data, the neonate MONSTR had few holes.

[^5]: Fun fact: Q1 is "universe" (appropriate) and Q2 is "Earth" (a little anthropocentric, humans).

## Cleaning

Throughout the data compiling and completing processes, we performed both targeted and random audits on SPARQL code performance and research assistants' outputs. Our greater burden is in justifying Wikipedia itself. It carries a strong stigma in scholarly research that will be difficult to shake [@becker_researchfauxpas_2015].[^6] We submit, however, that for certain types of information it is a useful source. There is virtually no scholar who promotes Wikipedia citations, but there is also virtually no scholar who does not queue up a Wiki page to get a basic sense or check a minor factoid on a topic. We continue to side with consensus that Wikipedia should not serve as an academic source for ambiguous, contested, or complex concepts and measures. It features several objective datapoints, though, that are amenable and appropriate to scholarly analysis. When collected carefully and used with intellectual honesty, we contend that the breadth, depth, and granularity of such information easily extracted through automated processes is worth the taboo trade-off.

[^6]: Please go read the entertaining opening of this short article, which we concede is better than ours.

Although the use of Wikipedia in social science research remains relatively nascent, much has been done to evaluate its validity and reliability [@greenstein_wikipediabiased_2012; @jullien_whatweknow_2012; @mesgari_sumallhuman_2015; @benjakob_clockworkwikipediabroad_2018]. In studies of over 10,000 randomly selected articles, scholars found Wikipedia's accuracy to be comparable to the Encyclopedia Britannica [@blumenstock_automaticallyassessingquality_2008], the Dictionary of American History, and American National Biography Online [@rector_comparisonwikipediaother_2008] using predictors such as page length, number of citations, external links, and edit history.[^7] Specific to political science, a study on Wikipedia's coverage of 246 major-party gubernatorial candidates contained no errors in candidate biographies and 242 out of the 246 (98.4%) reported election results within 0.2% of the actual outcome [@brown_wikipediadatasource_2011]. Validation exercises on party ideology data from Wikipedia similarly deem that it bears correct, consistent information [@munzert_measuringimportancepolitical_2018; @poschmann_disambiguatingspecifyingsocial_2022; @gobel_comparativelegislatorsdatabase_2021; @herrmann_partypositionswikipedia_2021].

[^7]: Of the nefarious last, studies find that contrary to many university course syllabi, Wikipedia vandalism is not as common as is often alleged: 81% of uncited claims are flagged by Wikipedia bots using automated vandalism detection within 24 hours of being posted and fixed within three hours [@adler_wikipediavandalismdetection_2011; @tramullas_researchwikipediavandalism_2016].

Inconsistencies largely pertain to frequently disputed figures (i.e., the precise size of armed forces in a battle) and more subjective interpretations (i.e., rationales and contested effects), datapoints that all sources struggle to pinpoint and discrepancies reflected across multiple venues. On fact-based claims (such as when and where an event occurred), Wikipedia is as accurate and often more comprehensive [@rector_comparisonwikipediaother_2008]. Since our dataset comprises objective claims, we proceed with some confidence. Still, in addition to front-end and process audits, once fully compiled the authors manually checked variable values for a randomly selected sample of 6% of observations. With the exception of uncertainty about the precise end date of two observations, Wikipedia's coding for all other values could be independently verified using external sources. Aware that readers might still be thinking we're monsters, our data are publicly available and we invite the most vicious of audits in the event it can be improved.

## Clustering

Existing datasets on military intervention exhibit a concerning assumption of independence across cases in a given sample. All valuable, respected, and oft used datasets -- International Crisis Behavior [@brecher_studycrisis_1997], UCDP/PRIO Armed Conflict Dataset [@gleditsch_armedconflict19462001_2002], Military Intervention by Powerful States (MIPS) [@sullivan_militaryinterventionpowerful_2009], International Military Intervention (IMI) [@pickering_internationalmilitaryintervention_2009], Militarized Interstate Dispute (MIDs) [@palmer_mid4dataset2002_2015] -- this assumption limits the ability to analyze decision-making and strategic behavior in the same dependent, dynamic context as political and military leaders. As an example, while ICB, MIDS, and MIPS have one observation for the two-decade War in Afghanistan (2001), IMI considers the attacking and peacekeeping portions of that intervention as separate and RAND includes Operation Enduring Freedom, Operation Swift Freedom, International Security Assistance Force, Resolute Support and Freedom's Sentinel. Kudos, RAND, but it gets thicker. In reality, Operation Swift Freedom was *part of* Operation Enduring Freedom - Afghanistan (the broader operation transcending the Afghan context as the "War on Terror"), Freedom's Sentinel was a sub-campaign of Resolute support, of course all of these occurred within the broader Afghanistan War rather than in isolation.

Wikipedia identifies which interventions are "part of" other interventions, enabling us to identify and code the operational and strategic levels in which each ought to be nested. These hierarchically nested intervention relationships are a unique feature of our dataset that yields several advantages. It allows us to code, for example, how many days into Resolute Support did Freedom's Sentinel occur. Since the unit of analysis is the military intervention subset to each "leaf" node, it facilitates hierarchical models accounting for structural dependencies among interventions borne of the same strategic campaign and parent-intervention fixed effects. Even in a bare-bones linear model with parent indicators, controlling for these relationships better reflects reality. Overall, despite its anathema origin, the more comprehensive, granular, and interdependent dataset we scraped together adds value to scholarship on the use of US military force. In the next sections, we aim to demonstrate this.

# Distinctive outputs

## Breadth of coverage

```{r breadth}
compare <- read.csv(paste0(here::here(), "/data/","02c_interventions_newdata_merged.csv")) %>%
  dplyr::select(wikidata_id, dplyr::starts_with("dataset_")) %>%
  dplyr::distinct() %>%
  dplyr::select(dplyr::starts_with("dataset_")) %>%
  dplyr::rename_at(vars(dplyr::starts_with("dataset_")),
                   dplyr::funs(stringr::str_replace(., "dataset_", ""))) %>%
  dplyr::rename(Allen = allen,
                Horowitz = horowitz,
                ICB = icb,
                MIPS = mips,
                IMI = imi,
                MID = mids,
                MIP = mip,
                RAND = rand) %>%
  replace(is.na(.), 0) %>%
  dplyr::mutate('MONSTR' = 1)
```

As already acknowledged, there are several datasets that feature military interventions: IMI [@pearson_internationalmilitaryintervention_1993; @kisangani_internationalmilitaryintervention_2008], ICB [@brecher_studycrisis_1997], UCDP PRIO [@gleditsch_armedconflict19462001_2002], MIDs [@palmer_mid4dataset2002_2015], newly MIP that builds upon MIDs [@kushi_introducingmilitaryintervention_2022], not to be confused with MIPS [@sullivan_militaryinterventionpowerful_2009], RAND [@kavanagh_characteristicssuccessfulmilitary_2019], the two excellent aerial bombing datasets by @horowitz_whendoesaerial_2001 and @allen_understandingimpactair_2017, and more.\^\[We assume that most readers who have made it this far are those for whom these acronyms are second nature. If not, they are defined in the "Clustering" subsection above.\] There is significant variation, however, in the cases they include. Of course, much of this is by design. Many were not devised to exclusively or completely cover military interventions. Some datasets take a broader approach in defining and identifying cases---international crises, militarized disputes---some narrower, examining a specific type of intervention. Nonetheless, some of the variation is due to constraints rather than design. When subset to our definition of US military interventions since 1989, the heterogeneity in coverage is stark. Only `r compare %>% dplyr::select(-MONSTR) %>% dplyr::mutate(sum = rowSums(.)) %>% dplyr::filter(sum == 8) %>% nrow()` case appears in all existing datasets. Prior to MIP's recent debut [@kushi_introducingmilitaryintervention_2022], the combined coverage of all these datasets post-1989 yields only `r compare %>% dplyr::select(-MONSTR) %>% dplyr::mutate(sum = rowSums(.)) %>% dplyr::filter(sum >= 1) %>% nrow()` unique interventions (MIP captures 37 TO DO:does MIP capture any new unique observations or are they all the same as MIDs / others? If not, we'll just include them in this sentence without distinction. Or raise the combined number a little if they got a couple more.). Our dataset provides information on `r compare %>% nrow()` US military interventions during this time period. @fig-comparison compares MONSTR to other datasets to demonstrate the large variation in coverage and its considerable expansion in cases within its time scope.

```{r}
#| label: fig-comparison
#| fig-cap: Comparison of dataset coverage. MONSTR includes every post-1991 US intervention in any existing dataset as well as 425 cases not included in any existing dataset.
#| fig-asp: 1
#| fig-width: 8

UpSetR::upset(compare,
      nsets = 9,
      number.angles = 0,
      point.size = 2.5,
      line.size = 1,
      text.scale = 1.8,
      mainbar.y.label = "Number of Interventions",
      sets.x.label = "Observations",
      set_size.numbers_size = 1.2,
      order.by = "freq")
```

We have already sketched the general reason for this - delving into the Wikiverse. The more specific reason is that Wiki enables us to disaggregate the unit of analysis to a more granular level. Current datasets typically identify and code variables associated with wars, such as the 2003 Iraq War. Ours identifies military operations and codes covariates associated with each individual event, such as Operation Viking Hammer conducted within the Iraq War. This disaggregation is imperative for understanding how states fight. @hart_strategyindirectapproach_1967 defines military strategy as "the art of distributing and applying military means to fulfill the ends of policy" (335). US Army Field Manual 100-5 (1993) affirms that strategic objectives and military battles are harmonized by *operational art*.[^8] Consequently, we see operations as the appropriate level for analyzing military strategy. Many scholars agree. For instance, @allen_understandingimpactair_2017 remark that "\[i\]deally, the unit of analysis for this project would be at the campaign level. Unfortunately, while most modern Western militaries operate in terms of campaigns, data for all countries over time was not available at the campaign level" (549). But like going to war with the Army you have rather than the one you want, scholars do the best they can with the data that exist. Stooping to Wikipedia, we can now go to war (with unanswered research questions) with the data we want.

[^8]: The @army_fm1005operations_1993 defines operational art as "the employment of military forces to attain strategic goals in a theater of war...through the design, organization, and conduct of campaigns and major operations" (6-2).

## Depth of datapoints

In addition to improving the granularity of our unit of analysis, we were able to extract more granular variables characterizing interventions. Shown in Table 1, MONSTR stacks up well against existing datasets on objective factors. It features start and end dates for each intervention down to the day and a count of casualties (including civilians) for each side. The dataset also contains indicators for seven different types of uses of force, a significant improvement over blunt domain codings or focus on a single type of force when it comes to our research questions. Most interventions include geocoded latitude and longitude coordinates, facilitating finer spatial analyses. It also identifies all state and nonstate actors active in the intervention, both intervenors and targets. Notably, MONSTR does not feature information on the objective and outcome of the military intervention. Treading lightly through data norms, we recognize that there is subjectivity afoot in discerning factors like military versus political objectives, *ex ante* and *ex post* objectives, and in interpreting outcomes.

```{r depth}
#| label: tbl-depth
#| tbl-cap: "Comparison of intervention-level covariates"

googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1XNkYLhErXdslRXtvpxv7jdz_ha2hjqq-rOOZGIfxw0Y/edit#gid=0",
                                    sheet = "compare",
                                    col_types = "c") %>%
  knitr::kable(format = "latex",
               escape = FALSE,
               align = rep('c', 11)) %>%
  kableExtra::row_spec(0,
                       bold = TRUE,
                       angle = 45) %>%
  kableExtra::landscape()
```

## Means of Force

To analyze military strategy, we remind of @hart_strategyindirectapproach_1967's seminal description as "the art of distributing and applying \textit{military means} to fulfill the ends of policy" (351, emphasis added). The dearth and incomparability of the means of force in existing datasets are chief reasons we considered condescending to Wikipedia in the first place. An attractive aspect of our dataset (to us) is that we could extract this variable for each intervention. After several iterations of coding, auditing, and refining, we observed a cooccurrence of seven main categories of force: ground troops, paramilitary, close air support, air to air combat, aerial bombing, cruise missiles, and drones. @tbl-means compares the means of force identified across existing datasets and ours. By our definition of intervention, we do not observe any naval interventions within our time scope. We do capture several nuanced means of land and air applications of force while other sources depict them in broad domains.

Each force structure has its trade offs. Boots on the ground are more versatile, discriminating in context, and effective at winning hearts and minds in counterinsurgent settings (when done well). At the same time, they are more directly exposed to enemy attack, implying higher casualties, and by numbers and interaction density alone more prone to natural human error. Manned aerial platforms significantly reduce risks and a degree of error, but exchange this for context-specific and time-sensitive intelligence and population-centric currency [@lyall_ragemachinesexplaining_2009]. Unmanned platforms (cruise missiles and drones) represent the lowest tactical risk to troops but the highest political risks of blowback [@fair_pakistanioppositionamerican_2014; @johnston_impactusdrone_2016] and downsides to decapitation campaigns [@price_targetingtopterrorists_2012; @gobeil_pickyourpoison_2017]. In addition to risks and effect, each of these means differently signals resolve [@zegart_cheapfightscredible_2018]. Overall, we assess that possessing the means of force for US military interventions grants us great purchase in the study of military strategy.

```{r means}
#| label: tbl-means
#| tbl-cap: "Comparison of type of use of force used in each US military intervention"

df_raw <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wiiuNVE2Wt-ZWyZRWDoqyxvg-y-hWpNdDWwdZjYCvM4/edit?usp=sharing",
                                    sheet = "04a_interventions_newdata_childof") %>%
  dplyr::select(wikidata_name, childof_name, realparent) %>%
  dplyr::filter(realparent == 1 |
                  is.na(realparent)) %>%
  dplyr::distinct() %>%
  dplyr::select(-realparent) 

df_parent <- df_raw %>%
  dplyr::select(childof_name) %>%
  dplyr::mutate(parent = 1)

df_child <- df_raw %>%
  dplyr::select(wikidata_name)

df_childonly <- dplyr::left_join(df_child, df_parent, by = c("wikidata_name" = "childof_name")) %>%
  dplyr::distinct() %>%
  dplyr::filter(is.na(parent))

counts <- read.csv(paste0(here::here(), "/data/","03a_intervention_newdata_dbpedia.csv"))

# Merge
df_childonly <- dplyr::left_join(df_childonly, counts) %>%
  dplyr::select(-parent)

df_childonly <- df_childonly %>% 
  dplyr::select(wikidata_name, dplyr::starts_with("means_")) %>%
  dplyr::rename("Aerial bombing" = means_aerialbombing, 
                "Air-to-air" = means_airtoair, 
                "Close Air Support" = means_closeairsupport, 
                "Cruise missiles" = means_cruisemissiles, 
                "Drones" = means_drones, 
                "Ground troops" = means_groundtroops, 
                "Paramilitary" = means_paramilitary) %>%
  dplyr::distinct() %>%
  tidyr::gather(wikidata_name, means) %>%
  dplyr::count(wikidata_name, means) %>%
  dplyr::filter(means == 1) %>%
  dplyr::select(-means) %>%
  dplyr::rename(means = wikidata_name) %>%
  dplyr::mutate(dataset = "wiki")

# Show a comparison with the other datasets that have means coded
imi <- imi %>%
  dplyr::mutate(Land = ifelse(type <= 3 & type >= 5, 1, 0),
                Sea = ifelse(naval == 4 | naval == 5, 1, 0),
                Air = ifelse(air == 3 | air == 4, 1, 0)) %>%
  dplyr::select(description, Land, Sea, Air) %>%
  tidyr::gather(description, means) %>%
  dplyr::count(description, means) %>%
  dplyr::filter(means == 1) %>%
  dplyr::select(-means) %>%
  dplyr::rename(means = description) %>%
  dplyr::mutate(dataset = "IMI")

mips <- mips %>%
  dplyr::mutate(Land = ifelse(forcetyp %in% c("ground combat", "limited ground"), 1, 0),
                Sea = ifelse(forcetyp == "naval", 1, 0),
                Air = ifelse(forcetyp == "extensive air", 1, 0)) %>%
  dplyr::select(case, Land, Sea, Air) %>%
  tidyr::gather(case, means) %>%
  dplyr::count(case, means) %>%
  dplyr::filter(means == 1) %>%
  dplyr::select(-means) %>%
  dplyr::rename(means = case) %>%
  dplyr::mutate(dataset = "MIPS")

mip <- mip %>%
  dplyr::mutate_at(c('MaxTroops', 'MaxNavy', 'MaxAir'), as.numeric) %>%
  dplyr::mutate(Land = ifelse(MaxTroops == 0 |
                                     is.na(MaxTroops), 0, 1),
                Sea = ifelse(MaxNavy == 0 |
                                   is.na(MaxNavy), 0, 1),
                Air = ifelse(MaxAir == 0 |
                                  is.na(MaxAir), 0, 1)) %>%
  dplyr::select(Name, Land, Sea, Air) %>%
  tidyr::gather(Name, means) %>%
  dplyr::count(Name, means) %>%
  dplyr::filter(means == 1) %>%
  dplyr::select(-means) %>%
  dplyr::rename(means = Name) %>%
  dplyr::mutate(dataset = "MIP")

allen_17 <- allen_17 %>%
  dplyr::mutate(Land = ifelse(othforces == 1, 1, 0),
                Sea = 0,
                Air = ifelse(bomb == 1, 1, 0)) %>%
  dplyr::select(war, Land, Sea, Air) %>%
  tidyr::gather(war, means) %>%
  dplyr::count(war, means) %>%
  dplyr::filter(means == 1) %>%
  dplyr::select(-means) %>%
  dplyr::rename(means = war) %>%
  dplyr::mutate(dataset = "Allen")

horowitz_01 <- horowitz_01 %>%
  dplyr::mutate(Land = 0,
                Sea = 0,
                Air = 1) %>%
  tibble::rowid_to_column("id") %>%
  dplyr::select(id, Land, Sea, Air) %>%
  tidyr::gather(id, means) %>%
  dplyr::count(id, means) %>%
  dplyr::filter(means == 1) %>%
  dplyr::select(-means) %>%
  dplyr::rename(means = id) %>%
  dplyr::mutate(dataset = "Horowitz")

# Merge all into a single table
means_tab <- dplyr::bind_rows(df_childonly, imi, mips, mip, allen_17, horowitz_01) %>%
  tidyr::pivot_wider(names_from = dataset,
                     values_from = n) %>%
  dplyr::mutate(means = dplyr::recode(means,
                                      "Land" = "All land",
                                      "Sea" = "All naval",
                                      "Air" = "All air"),
                means = as.factor(means),
                means = forcats::fct_relevel(means, c('All air', 'Aerial bombing', 'Air-to-air', 'Close Air Support', 'Cruise missiles', 'Drones', 'All naval', 'All land', 'Ground troops', 'Paramilitary')))

# Final table
means_tab %>%
  dplyr::arrange(means) %>%
  gt::gt(rowname_col = "means") %>%
  gt::tab_header(title = "Comparison of Datasets Identifying Military Strategies",
                 subtitle = "United States (1991-2021)") %>%
  gt::tab_row_group(label = "Sea",
                    rows = 7) %>%
  gt::tab_row_group(label = "Land",
                    rows = 8:10) %>%
  gt::tab_row_group(label = "Air",
                    rows = 1:6) %>%
  gt::sub_missing(columns = everything(),
                  missing_text = "-") 
# %>% gt::gtsave("compare.tex")
```

## Nested Interdependencies

Researchers have long expressed concern over un-modeled interdependencies between conflict events. While frequently acknowledged in theories, the empirical mismatch due to limitations in available data problematize the important inferences scholars can make. Leveraging the structure of Wikipedia's internal architecture, we identified and coded the nested relationships for each intervention. @fig-dendrogram illustrates the high degree of interdependency among interventions in our dataset. @fig-nested hones in on the 1991 Gulf War to provide a more specific example. The Gulf War can be decomposed into three broad approaches: the no-fly zones, liberation of Kuwait, and air campaigns. Each features several individual military operations in support of its objective. Looking at the distribution of the means of force along the right side of the figure, the importance of controlling for strategic campaigns becomes apparent. Intuitively, the aerial-focused ones feature higher frequencies of air power while the liberation campaign involves more ground troops, often in tandem with close air support. Although Wikipedia's nested dependencies are not a perfect mapping relative to military or political dependencies, they nonetheless will help scholars to better model this vital trait. In addition, they enable us to look beyond the onset of war to examine and explain the evolution of strategy over its course.

```{r}
#| label: fig-dendrogram
#| fig-cap: "Complete nested dependencies of US interventions. Each intervention is a 'part of' the intervention connected to its left."
#| fig-height: 10

# Edge list
dependency <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wiiuNVE2Wt-ZWyZRWDoqyxvg-y-hWpNdDWwdZjYCvM4/edit?usp=sharing",
                                    sheet = "04a_interventions_newdata_childof") %>%
  dplyr::select(wikidata_name, childof_name, realparent) %>%
  dplyr::filter(realparent == 1 |
                  is.na(realparent)) %>%
  dplyr::distinct() %>%
  dplyr::select(-realparent) %>%
  dplyr::mutate(childof_name = dplyr::if_else(childof_name == "NA", "intervention", childof_name)) %>%
  dplyr::rename(child = wikidata_name,
                parent = childof_name) %>%
  dplyr::select(parent, child) %>%
  dplyr::mutate_all( ~ stringr::str_replace_all(., " ", "_")) %>%
  na.omit()

# Testing
library(igraph)
tree <- igraph::graph_from_edgelist(as.matrix(dependency))

library(ggraph)
ggraph(tree, layout = 'tree') +
  geom_edge_diagonal(edge_width = 0.05,
                 color = 'steelblue') +
  geom_node_point(size = 0.05) +
  coord_flip() +
  scale_y_reverse() +
  theme_void()
```

```{r nested}
#| label: fig-nested
#| fig.cap: "Nested structure of Gulf War (1991). The war was divided into roughly 3 broader campaigns, within which individual operations were nested. Red cells indicate the US used the corresponding military strategy during that operation. Gray cells indicate use unknown."
#| fig.height: 4
#| fig.width: 6

# Edge list
dependency <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wiiuNVE2Wt-ZWyZRWDoqyxvg-y-hWpNdDWwdZjYCvM4/edit?usp=sharing",
                                    sheet = "04a_interventions_newdata_childof") %>%
  dplyr::select(wikidata_name, childof_name, realparent) %>%
  dplyr::filter(realparent == 1 |
                  is.na(realparent)) %>%
  dplyr::distinct() %>%
  dplyr::select(-realparent) %>%
  dplyr::mutate(childof_name = dplyr::if_else(childof_name == "NA", 
                                              "intervention", childof_name)) %>%
  dplyr::rename(child = wikidata_name,
                parent = childof_name) %>%
  dplyr::select(parent, child) %>%
  dplyr::mutate_all( ~ stringr::str_replace_all(., " ", "_")) %>%
  na.omit() %>%
  dplyr::filter(parent %in% c("Gulf_War",
                              "Gulf_War_air_campaign",
                              "Iraqi_no-fly_zones",
                              "Liberation_of_Kuwait_campaign",
                              "Operation_Desert_Shield",
                              "Lead-up_to_the_Iraq_War",
                              "Operation_Southern_Watch",
                              "Iraq_disarmament_crisis")) %>%
  dplyr::filter(!child %in% c("Operation_Provide_Comfort",
                              "Operation_Instant_Thunder",
                              "Operation_Desert_Thunder",
                              "Operation_Desert_Shield",
                              "Kurdistan_Islamist_Conflict",
                              "Defense_of_Saudi_Arabian_country",
                              "Battle_of_Bubiyan",
                              "Iraqi_Kurdish_Civil_War"))

# Defense of Saudi Arabian country for the period 2 August 1990, through 16 January 1991.
# Liberation and Defense of Kuwait for the period 17 January 1991, through 11 April 1991.
# Southwest Asia Cease-Fire for the period 12 April 1991, through 30 November 1995, including Operation Provide Comfort.

# Phylo tree
library(ggtree)
tree <- dependency %>%
  treeio::as.phylo() %>%
  ggtree::ggtree(color="black",
                 linetype = 1,
                 layout = "roundrect") +
  geom_tiplab(align = TRUE,
              size = 3) +
  geom_nodelab(geom = "label",
               size = 3) +
  xlim(-0.1, 4.5)

tree$data$label <- stringr::str_replace_all(tree$data$label, "_", " ")

# Heatmap of the above child nodes
heat <- dependency %>%
  dplyr::filter(child %in% c("Amiriyah_shelter_bombing",
                             "Attack_on_Ras_Tanura",
                             "Battle_of_Ad-Dawrah",
                             "Battle_of_Khafji",
                             "Battle_of_Qurah_and_Umm_al_Maradim",
                             "Battle_of_Wadi_Al-Batin",
                             "Battle_for_Jalibah_Airfield",
                             "Battle_of_73_Easting",
                             "Battle_of_Al_Busayyah",
                             "Battle_of_Kuwait_International_Airport",
                             "Battle_of_Medina_Ridge",
                             "Battle_of_Norfolk",
                             "Battle_of_Phase_Line_Bullet",
                             "Battle_of_Rumaila",
                             "Package_Q_Strike",
                             "Samurra_Air_Battle",
                             "Cruise_missile_strikes_on_Iraq",
                             "1993_cruise_missile_strikes_on_Iraq",
                             "Bombing_of_Iraq",
                             "January_1993_airstrikes_on_Iraq",
                             "Operation_Southern_Focus",
                             "Operation_Vigilant_Warrior",
                             "Operation_Vigilant_Sentinel",
                             "Operation_Desert_Spring",
                             "Operation_Northern_Watch",
                             "Operation_Senior_Surprise"))

df_interv_heat <- df_interv %>%
  dplyr::select(wikidata_name, dplyr::starts_with("means_")) %>%
  dplyr::mutate_all( ~ stringr::str_replace_all(., " ", "_"))

heat <- dplyr::left_join(heat, df_interv_heat, by = c("child" = "wikidata_name")) %>%
  dplyr::select(child, dplyr::starts_with("means_"))  %>%
  tidyr::pivot_longer(cols = dplyr::starts_with("means_"),
                      names_to = "means",
                      values_to = "used") %>%
  dplyr::mutate(means = stringr::str_remove(means, "means_"),
                used = as.numeric(used))

heat <- heat %>%
  dplyr::mutate(child = stringr::str_replace_all(child, "_", " ")) %>%
  dplyr::mutate(child = as.factor(child)) %>%
  dplyr::mutate(child = forcats::fct_relevel(child, 
                                             levels = c("Amiriyah shelter bombing",
                                                        "Attack on Ras Tanura",
                                                        "Package Q Strike",
                                                        "Samurra Air Battle",
                                                        "Cruise missile strikes on Iraq",
                                                        "Operation Vigilant Sentinel",
                                                        "Operation Desert Spring",
                                                        "Operation Northern Watch",
                                                        "1993 cruise missile strikes on Iraq",
                                                        "Bombing of Iraq",
                                                        "January 1993 airstrikes on Iraq",
                                                        "Operation Southern Focus",
                                                        "Operation Vigilant Warrior",
                                                        "Battle for Jalibah Airfield",
                                                        "Battle of 73 Easting",
                                                        "Battle of Ad-Dawrah",
                                                        "Battle of Al Busayyah",
                                                        "Battle of Khafji",
                                                        "Battle of Kuwait International Airport",
                                                        "Battle of Medina Ridge",
                                                        "Battle of Norfolk",
                                                        "Battle of Phase Line Bullet",
                                                        "Battle of Qurah and Umm al Maradim",
                                                        "Battle of Rumaila",
                                                        "Battle of Wadi Al-Batin",
                                                        "Operation Senior Surprise")))

hmap <- ggplot2::ggplot(heat,
                  aes(y = child, x = means, fill = used)) +
  geom_tile(color = "white",
            lwd = 0.3) +
  scale_fill_gradient(low = "white", high = "red") +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 90,
                                   vjust = -1),
        axis.title.x = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  coord_fixed() +
  scale_x_discrete(position = "top",
                   limits = c("drones",
                              "cruisemissiles",
                              "airtoair",
                              "aerialbombing",
                              "closeairsupport",
                              "groundtroops",
                              "paramilitary"),
                   labels = c("Drones",
                              "Cruise Missiles",
                              "Air-to-Air",
                              "Aerial Bombing",
                              "Close Air Support",
                              "Ground Troops",
                              "Paramilitary"))

cowplot::plot_grid(tree, hmap, align = "h", rel_widths = c(4.7, 1))
```

## Geographic Information

Nearly all interventions in MONSTR are geocoded with GPS coordinates indicating where the intervention occurred. In @fig-map, we plot the location of the US interventions in the dataset. For spatial analysts and geographic information system (GIS) users, this opens a wealth of inquiries. For instance, it allows for the coding of terrain, which might influence the force structure deployed [@gleditsch_richardsoninformationage_2012; @shaver_terrainruggednessland_2019]. It also allows analysis of urbanicity, proximity to strategic assets, intervention amid shifting combat fronts, power projection, contagion, diffusion, and more (limited only by GIS layers and imagination). As a GIS-capable team, we are enthusiastic about these prospects. Still, as an intellectually honest team, we caution interpretations about the geography of US military interventions from these data for two reasons. First, Wikipedia does not distinguish between geometry type, which matters for interventions that occurred in multiple districts or towns. Second, we addressed nonrandom missingness using Google Earth data, which raises questions about the consequences of geographic coding using multiple gazetteers [@douglass_measuringlandscapecivil_2018].

```{r map}
#| label: fig-map
#| fig.cap: "Caption for map TBD."
#| fig.height: 5
#| fig.width: 6

df_sf <- readRDS(paste0(here::here(), "/data/","06a_model.rds")) %>%
  dplyr::filter(!is.na(lat) & !is.na(lon)) %>%
  sf::st_as_sf(coords = c("lon", "lat"))

# Calculate center for view
WorldData <- rworldmap::getMap(resolution = "high") %>%
  sf::st_as_sf()

theme_map <- function(base_size = 9, base_family = "") {
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          panel.background = element_blank(),
          panel.border = element_blank(),
          panel.grid = element_blank(),
          panel.spacing = unit(0, "lines"),
          plot.background = element_blank(),
          legend.justification = c(0, 0)#,
          #legend.position = c(1,1)
    )
  }

df_map <- df_interv %>%
  dplyr::filter(!is.na(urban_binary)) %>%
  dplyr::mutate(means_ground = ifelse(means_groundtroops == 1 |
                                        means_paramilitary == 1, 1, 0),
                means_air = ifelse(means_aerialbombing == 1 |
                                     means_airtoair == 1 |
                                     means_drones == 1 |
                                     means_closeairsupport == 1 |
                                     means_cruisemissiles == 1, 1, 0),
                means_factor = ifelse(means_ground == 1 &
                                        means_air == 0, "Ground", NA),
                means_factor = ifelse(means_air == 1 &
                                        means_ground == 0, "Air", means_factor),
                means_factor = ifelse(means_ground == 1 &
                                        means_air == 1, "Mixed", means_factor),
                urban_binary = ifelse(urban_binary == 1, "Urban Target", "Non-urban Target")) %>%
  dplyr::filter(!is.na(means_factor))

ggplot(data = WorldData) +
  geom_sf(fill = "antiquewhite1",
          size = 0.25) +
  coord_sf(xlim = c(35, 75), ylim = c(28, 39), expand = FALSE) +
  geom_sf_text(data = WorldData[WorldData$SOVEREIGNT %in% c("Afghanistan", "Pakistan", "Iraq", "Syria", "Kuwait"), ],
               aes(label = NAME),
               size = 2.5) +
  geom_point(data = df_map,
             aes(x = lon, y = lat,
                 color = means_factor),
             size = 0.8) +
  facet_wrap(~ urban_binary,
             ncol = 1) +
  scale_color_manual(values = c("#7570b3", "#1b9e77", "#d95f02")) +
  xlab("") + ylab("") + labs(color = "Military Means") +
  theme(panel.grid.major = element_line(color = gray(.5),
                                        linetype = "dashed", size = 0.25),
        panel.background = element_rect(fill = "aliceblue"),
        legend.position = "bottom",
        axis.text = element_text(size = 5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(hjust = 0),
        plot.margin = margin(0, 0, 0, 0, "cm")) +
  guides(color = guide_legend(override.aes = list(size = 4)))
```

## Actors Involved

The identity, traits, and connections of actors in conflict have a dramatic impact on how they fight. Scholars have long recognized and researched this [@mack_whybignations_1975; @huth_majorpowerintervention_1998]. Powerful states, more likely to have economic assets to protect and military ones to project power, intervene more frequently [@lake_powerfulpacifistsdemocratic_1992; @sullivan_militaryinterventionpowerful_2009]. Democracies might deploy force for different reasons, in different ways, and for different durations than autocracies [@caverley_democraticmilitarismvoting_2014; @talmadge_differentthreatsdifferent_2016]. Some states are primarily compelled by security, some economy, some to contain spillover of violence or refugees, some drawn in by allies or against rivals, and some by humanitarian impulses. Knit by political, economic, and cultural legacies, former colonial relations increase the probably of intervention [@chacha_colonialtiescivil_2019]. Considering intervenors alone, multilateral campaigns---distributing costs and risks across partners yet involving larger coordination challenges---require different calculus than unilateral ones. For example, in Serbia 1996 US generals realized that to keep the 19 NATO allies on board, their preference of a quick, decisive air campaign was off the table [@cooper_politicsairstrikes_2001]. Speaking only of targets, states might behave differently against nonstate combatants relative to fellow states. And in the same way that multilateral state coalitions change the fighting calculus, nonstate actors form complex and shifting coalitions too. Just look at the Syria Civil War, with its makeshift operations rooms and mercurial alliances [@cranmer_coalitionqualitymultinational_2018]. Furthermore, the consequences of intervention hinge on collective actor traits on both sides and interactions across [@carment_threecompanyevaluating_1998; @caverley_mythmilitarymyopia_2009; @chiba_militarycoalitionscrisis_2019].

While we know these factors matter, much of the analysis focusing on them oversimplifies by looking at the primary intervenor and target. This masks important variation in participants, especially those occurring over years or even decades. The roster comprising the coalition throughout the Afghanistan War, for instance, oscillated over its life. Even if it remained constant, not all members participated in each operation [@gannon_keepingyourfriends_2021]. Nor did the targets remain constant across both decades, across every operation. At times, the US unilaterally deployed special forces against anti-coalition insurgents (Operation Red Wings). At other times, it pitched in beside Dutch pilots to provide close air support for Aussies and their Afghan mentees ambushed by the Taliban (First Battle of Kakarak). Unsurprisingly, the force structures selected for these events were quite different. Who fights, whom they are fighting alongside, and whom they are fighting against are important for explaining comportment in conflict [@cappellazielinski_organizingperformancecoalition_2020].

This variation in actor interaction is showcased in three network plots of the US and its alliance partners (side A) in Figure 6. Nodes labeled in green are state actors, those in orange nonstate groups. The edges are weighted by the number of operations shared between dyads. Maintaining focus on the Afghanistan War, the top right panel portrays the many actors participating throughout the 20-year campaign. Intuitively, the US most densely interacted with the state of Afghanistan itself after ousting the Taliban. While key Western partners situate strongly in the network, direct US involvement with the UK was most prominent. Other interesting traits emerge in comparison to the 1991 Gulf War (top left panel) and the 2015 Syria Civil War (bottom left). First, the composition of US partners changed dramatically across these conflicts. All allies in the Gulf War were states. In Afghanistan, state partners remain the norm but plenty of nonstate actors joined the fray. In the Syria War, nonstate partners outnumber states two-to-one. Edges shared with the UK are thick throughout, but in Syria US cooperation with a nonstate actor eclipses it by a small margin. Interaction density between actors varies across campaigns as well, being straightforward and balanced in the Gulf, layered and feathery in Afghanistan, and sprawling and asymmetric in Syria. Even with only allies in only three strategic events, we see intriguing patterns that spark questions. MONSTR affords the ability to analyze them in significantly more detail.

```{r actors}
#| label: fig-actors
#| fig.cap: "Caption TBD."

# load
parents <- readRDS(paste0(here::here(), "/data/","06a_model.rds")) %>%
  dplyr::select(wikidata_id, parent_id, parent_name)

# Gulf War
## New df of node attributes
nodes <- dplyr::left_join(counts, parents) %>%
  dplyr::filter(parent_name == "Gulf War") %>%
  dplyr::select(wikidata_name, belligerent_sideA) %>%
  dplyr::count(belligerent_sideA)

qcodes <- googlesheets4::read_sheet(ss = "https://docs.google.com/spreadsheets/d/1wiiuNVE2Wt-ZWyZRWDoqyxvg-y-hWpNdDWwdZjYCvM4/edit#gid=1180615821",
                                    sheet = "belligerent_list") %>%
  dplyr::filter(is.na(ccode_wrong)) %>%
  dplyr::select(belligerent, ccode, cname, Qcode) %>%
  dplyr::rename(belligerent2 = belligerent,
                ccode2 = ccode,
                cname2 = cname,
                qcode2 = Qcode)

nodes <- dplyr::left_join(nodes, qcodes,
                          by = c("belligerent_sideA" = "belligerent2")) %>%
  dplyr::mutate(sideA_nonstate = dplyr::if_else(is.na(ccode2), 1, 0),
                label = dplyr::if_else(is.na(cname2), belligerent_sideA, cname2)) %>%
  dplyr::select(label, belligerent_sideA, n, sideA_nonstate) %>%
  dplyr::filter(!label %in% c("US-led coalition"))

## Prep adjacency matrix
actors <- dplyr::left_join(counts, parents) %>%
  dplyr::filter(parent_name == "Gulf War") %>%
  dplyr::select(wikidata_name, wikidata_id, belligerent_sideA) %>%
  dplyr::distinct()

net_actor <- dplyr::left_join(actors, nodes) %>%
  dplyr::select(wikidata_name, label) %>%
  dplyr::filter(!label %in% c("US-led coalition"),
                !is.na(label))

## Prep network
library(igraph)

network_gulf <- as.matrix(reshape2::dcast(wikidata_name ~ label, data = net_actor, length)[,-1]>0 )
network_gulf <- t(network_gulf) %*% network_gulf
network_gulf <- igraph::graph_from_adjacency_matrix(network_gulf,
                                               mode = "undirected",
                                               weighted = TRUE,
                                               diag = FALSE)

## Set node colors for state vs non-state
library(RColorBrewer)
colrs  <- RColorBrewer::brewer.pal(3, "Dark2") 
network_gulf <- set_vertex_attr(network_gulf, "type", nodes$label, nodes$sideA_nonstate)
V(network_gulf)["United States"]$type <- "center"
my_color_gulf <- colrs[as.numeric(as.factor(V(network_gulf)$type))]
 
deg <- degree(network_gulf, mode = "all")
E(network_gulf)$width <- E(network_gulf)$weight

## Visualize network
l <- layout_with_graphopt(network_gulf,
                          charge = 5)

# Afghanistan war
## New df of node attributes
nodes <- dplyr::left_join(counts, parents) %>%
  dplyr::filter(parent_name == "War in Afghanistan") %>%
  dplyr::select(wikidata_name, belligerent_sideA) %>%
  dplyr::count(belligerent_sideA)

qcodes <- googlesheets4::read_sheet(ss = "https://docs.google.com/spreadsheets/d/1wiiuNVE2Wt-ZWyZRWDoqyxvg-y-hWpNdDWwdZjYCvM4/edit#gid=1180615821",
                                    sheet = "belligerent_list") %>%
  dplyr::filter(is.na(ccode_wrong)) %>%
  dplyr::select(belligerent, ccode, cname, Qcode) %>%
  dplyr::rename(belligerent2 = belligerent,
                ccode2 = ccode,
                cname2 = cname,
                qcode2 = Qcode)

nodes <- dplyr::left_join(nodes, qcodes,
                          by = c("belligerent_sideA" = "belligerent2")) %>%
  dplyr::mutate(sideA_nonstate = dplyr::if_else(is.na(ccode2), 1, 0),
                label = dplyr::if_else(is.na(cname2), belligerent_sideA, cname2)) %>%
  dplyr::select(label, belligerent_sideA, n, sideA_nonstate) %>%
  dplyr::filter(!label %in% c("Coalition forces",
                              "International Security Assistance Force",
                              "ISAF",
                              "NATO",
                              "Resolute Support Mission"))

## Prep adjacency matrix
actors <- dplyr::left_join(counts, parents) %>%
  dplyr::filter(parent_name == "War in Afghanistan") %>%
  dplyr::select(wikidata_name, wikidata_id, belligerent_sideA) %>%
  dplyr::distinct()

net_actor <- dplyr::left_join(actors, nodes) %>%
  dplyr::select(wikidata_name, label) %>%
  dplyr::filter(!label %in% c("Coalition forces",
                              "International Security Assistance Force",
                              "ISAF",
                              "NATO",
                              "Resolute Support Mission"),
                !is.na(label))

## Prep network
library(igraph)

network_isaf <- as.matrix(reshape2::dcast(wikidata_name ~ label, data = net_actor, length)[,-1]>0 )
network_isaf <- t(network_isaf) %*% network_isaf
network_isaf <- igraph::graph_from_adjacency_matrix(network_isaf,
                                               mode = "undirected",
                                               weighted = TRUE,
                                               diag = FALSE)

# Set node colors for state vs non-state
library(RColorBrewer)
colrs  <- RColorBrewer::brewer.pal(3, "Dark2") 
network_isaf <- set_vertex_attr(network_isaf, "type", nodes$label, nodes$sideA_nonstate)
V(network_isaf)["United States"]$type <- "center"
my_color_isaf <- colrs[as.numeric(as.factor(V(network_isaf)$type))]
 
deg <- degree(network_isaf, mode = "all")
E(network_isaf)$width <- E(network_isaf)$weight/3

# Visualize network
l <- layout_with_graphopt(network_isaf,
                          charge = 5)

# Syrian War
## Prep
nodes <- dplyr::left_join(counts, parents) %>%
  dplyr::filter(parent_name == "Syrian Civil War") %>%
  dplyr::select(wikidata_name, belligerent_sideA) %>%
  dplyr::count(belligerent_sideA)

qcodes <- googlesheets4::read_sheet(ss = "https://docs.google.com/spreadsheets/d/1wiiuNVE2Wt-ZWyZRWDoqyxvg-y-hWpNdDWwdZjYCvM4/edit#gid=1180615821",
                                    sheet = "belligerent_list") %>%
  dplyr::filter(is.na(ccode_wrong)) %>%
  dplyr::select(belligerent, ccode, cname, Qcode) %>%
  dplyr::rename(belligerent2 = belligerent,
                ccode2 = ccode,
                cname2 = cname,
                qcode2 = Qcode)

sort(unique(nodes$label))

nodes <- dplyr::left_join(nodes, qcodes,
                          by = c("belligerent_sideA" = "belligerent2")) %>%
  dplyr::mutate(sideA_nonstate = dplyr::if_else(is.na(ccode2), 1, 0),
                label = dplyr::if_else(is.na(cname2), belligerent_sideA, cname2)) %>%
  dplyr::select(label, belligerent_sideA, n, sideA_nonstate) %>%
  dplyr::filter(!label %in% c("CJTF-OIR",
                              "Islamic State of Iraq and the Levant"))

# Prep adjacency matrix
actors <- dplyr::left_join(counts, parents) %>%
  dplyr::filter(parent_name == "Syrian Civil War") %>%
  dplyr::select(wikidata_name, wikidata_id, belligerent_sideA) %>%
  dplyr::distinct()

net_actor <- dplyr::left_join(actors, nodes) %>%
  dplyr::select(wikidata_name, label) %>%
  dplyr::filter(!label %in% c("CJTF-OIR",
                              "Islamic State of Iraq and the Levant"),
                !is.na(label))

# Prep network
library(igraph)

network_syria <- as.matrix(reshape2::dcast(wikidata_name ~ label, data = net_actor, length)[,-1]>0 )
network_syria <- t(network_syria) %*% network_syria
network_syria <- igraph::graph_from_adjacency_matrix(network_syria,
                                               mode = "undirected",
                                               weighted = TRUE,
                                               diag = FALSE)

# Set node colors for state vs non-state
library(RColorBrewer)
colrs  <- RColorBrewer::brewer.pal(3, "Dark2") 
network_syria <- set_vertex_attr(network_syria, "type", nodes$label, nodes$sideA_nonstate)
V(network_syria)["United States"]$type <- "center"
my_color_syria <- colrs[as.numeric(as.factor(V(network_syria)$type))]

deg <- degree(network_syria, mode = "all")
E(network_syria)$width <- E(network_syria)$weight

# Combine all plots
## Fix color matches by switching orange to purple for ISAF
my_color_gulf[my_color_gulf == "#D95F02"] <- "#7570B3"

par(mfrow = c(2, 2),
    mar = c(1, 0, 1, 0))

l <- layout_with_gem(network_gulf)
n_gulf <- plot(network_gulf,
               edge.color = "gray",
               vertex.shape = "none",
               vertex.label.cex = 0.7,
               vertex.label.color = my_color_gulf,
               layout = l,
               main = "Gulf War (1991)",
               asp = 0)

l <- layout_with_gem(network_isaf)
n_isaf <- plot(network_isaf,
               edge.color = "gray",
               vertex.shape = "none",
               vertex.label.cex = 0.7,
               vertex.label.color = my_color_isaf,
               layout = l,
               main = "War in Afghanistan (2001)",
               asp = 0)

l <- layout_with_gem(network_syria)
n_syria <- plot(network_syria,
                edge.color = "gray",
                vertex.shape = "none",
                vertex.label.cex = 0.7,
                vertex.label.color = my_color_syria,
                layout = l,
                main = "Syrian Civil War (2015)",
                asp = 0)

legend(x = 2.1, y = 0.6,
       c("State ally","Non-state ally", "US"),
       xpd = NA,
       pch = 21,
       col = "#777777",
       pt.bg = colrs, pt.cex = 2, cex = .8, bty = "n", ncol = 1)
```

# Demonstration

As a proof of concept illustrating how these data can extend our theoretical understanding of military interventions, this section provides an inductive machine-learning model that predicts the type of use of force the US will use in a given intervention. Here, we take advantage of new information disaggregating the dependent variable into types of use of force as well as fine-grained information in previously theorized independent variables concerning factors like geography, timing, coalition effort, target type, and incumbent.

@tbl-results shows the relationship between the type of use of force in a US military intervention and various covariates that existing research has theorized has a relationship with these interventions. This paper's innovation here is two-fold. First, we disaggregated the type of use of force to include aerial bombing and ground operations, as done by X, but also distinguish between the type of aerial bombing given research about their difference in cost and efficacy. Second, our model accounts for the non-independence of observations in being a multi-level model that adjusts for the fact that interventions that took place within the same larger war should have that similarity accounted for.

```{r}
#| label: tbl-results
#| tbl-cap: Regression results

# Prep vars
df_interv <- df_interv %>%
  dplyr::mutate(year_base = year - 1989,
                year_square = year_base^2,
                year_cube = year_base^3,
                duration_log = log(duration + 1),
                daysinto_parent_log = log(daysinto_parent + 1),
                presidentid = as.factor(presidentid),
                presidentid = relevel(presidentid, ref = "George Bush"),
                means_ground = ifelse(means_groundtroops == 1 |
                                        means_paramilitary == 1, 1, 0),
                means_air_manned = ifelse(means_aerialbombing == 1 |
                                     means_closeairsupport == 1, 1, 0),
                means_air_unmanned = ifelse(means_drones == 1 |
                                          means_cruisemissiles == 1, 1, 0),
                means_air_all = ifelse(means_air_manned == 1 |
                                          means_air_unmanned == 1, 1, 0),
                means_mixed = ifelse(means_ground == 1 &
                                          means_air_all == 1, 1, 0))

# Models
models <- list()

models[['Ground']] <- lme4::glmer(means_ground ~
                    sideB_nonstate_all_binary +
                    multilateral1 +
                    mountainous +
                    urban_binary +
                    duration_log +
                    daysinto_parent_log +
                    presidentid +
                    year_base +
                    year_square +
                    year_cube +
                    (1 | parent_id),
                  data = df_interv,
                  family = binomial,
                  control = lme4::glmerControl(optimizer = "bobyqa",
                                               optCtrl = list(maxfun = 2e5)),
                  nAGQ = 0)

models[['Ground and Air']] <- update(models[['Ground']], means_mixed ~ .)
models[['Air (all)']] <- update(models[['Ground']], means_air_all ~ .)
models[['Air (Piloted)']] <- update(models[['Ground']], means_air_manned ~ .)
models[['Air (Unpiloted)']] <- update(models[['Ground']], means_air_unmanned ~ .)

# Table of results
modelsummary::modelsummary(models,
                           statistic = 'p.value',
                           stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
                           coef_omit = "Intercept|presidentid*|year_",
                           coef_rename = c("daysinto_parent_log" = "Day into Parent (log)",
                                           "duration_log" = "Duration (log)",
                                           "urban_binary1" = "Urban target",
                                           "mountainous1" = "Mountainous target",
                                           "sideB_nonstate_all_binary1" = "Non-state target",
                                           "multilateral11" = "Multilateral"),
                           notes = list('All models include president fixed effects and cubic spline polynomials.'))

df_interv %>%
  dplyr::select(means_drones, means_airtoair, means_cruisemissiles, means_closeairsupport, means_groundtroops, means_paramilitary,
                duration, daysinto_parent,
                allycount1,
                sideB_state_count, sideB_nonstate_all_count,
                urban_binary, mountainous) %>%
  dplyr::mutate(duration = as.numeric(duration),
                sideB_state_count = as.numeric(sideB_state_count)) %>%
  stargazer::stargazer(title = "Covariate Summary Statistics",
                     covariate.labels = c("Drones", "Air to air", "Cruise missiles", "Close air support", "Ground troops", "Paramilitary",
                                          "Duration (days)", "Days into parent (days)",
                                          "Ally count", "Adversary count (state)", "Adversary count (non-state)", "Urban target", "Mountainous target"),
                     digits = 2,
                     notes = "All military means in the origina dataset shown. Their aggregation in the statistical model is discussed above.",
                     header = FALSE,
                     label = "tab:summary-stats",
                     type = "latex")
  
```

# Conclusion

We have scantly demonstrated the potential applications of these data, showing predictive trends by US president, actors involved, and some temporal-spatial factors. Some findings are intuitive, some provoking. We see many avenues for future research, both theoretical and empirical. %theoretical avenues here

Empirically, despite the considerable expansion in coverage and granularity, our data are bounded in a brief time period for one nation. Data stretching further back and abroad are there, awaiting extraction. Given the machine-readable format, it is exceptionally easier to approach an endeavor of this scope than with manual efforts. We have no expectation or hope that a Wiki-based dataset will replace or compete with others. We do, however, expect and hope that it will expand in coverage and use. We continue to acknowledge Wikipedia's drawbacks and limitations. Nonetheless, we maintain that, with intellectual honesty about these traits, we can learn a great deal about the conduct of war that we would otherwise not know. Combined with more rigorous qualitative and quantitative approaches, we believe our MONSTR makes a meaningful contribution to the social scientific and practitioner communities aiming to accumulate knowledge. We hope the MONSTR grows. Many will choose not to use it but we hope in the end some will profess, "I love you too much to condemn you" [@dracula_].

\newpage

# References
